### 2015-05-03 abstract.tex
stuff

\selectlanguage{english}
% \renewcommand{\abstractname}{\Large{The \lego Motorcyclist robot}}
\addcontentsline{toc}{chapter}{\numberline{}Abstract}
\begin{abstract}
	\begin{verse}
	\small \textit{
	by Quentin Delhaye, Arnaud Dumont, Camille Giaux, Pierre Lasbleis and  Gauthier Roig; Université Libre de Bruxelles, 2011–2012.}
	\normalsize
	\end{verse}
\end{abstract}

\clearpage

\selectlanguage{francais}
% \renewcommand{\abstractname}{\Large{Le robot motard \lego}}
\addcontentsline{toc}{chapter}{\numberline{}Résumé}
\begin{abstract}
	\begin{verse}
	\small \textit{
	par Quentin Delhaye, Arnaud Dumont, Camille Giaux, Pierre Lasbleis et Gauthier Roig ; Université Libre de Bruxelles, 2011–2012.}
	\normalsize
	\end{verse}
\end{abstract}