### 2015-05-03 main.tex

\documentclass[a4paper, 12pt, oneside]{memoir}

\input{preamble}

\begin{document}

\frontmatter

\include{PageDeGarde}

\include{abstract}

\include{aknowledgements}

\selectlanguage{english}
\tableofcontents \pagebreak

\mainmatter

\appendix

\backmatter

% Bibliography

\end{document}