### 2015-05-03 aknowledgements.tex
stuff

\addcontentsline{toc}{chapter}{\numberline{}Aknowledgements}
\chapter*{Aknowledgements}

Thanks a bunch of people here:
- Frédéric Robert for his support
- Sébastien Rabou for his insight
- Batien Heneffe for his extensive help