### 2015-05-03 presentation.tex

\chapter{Presentation}
Here we talk about the protocols, the platform.
Show The OS stack (kernel/user)

\section{Experimental setup}
The experimental environment is build around a standard x86 host and an ARM Cortex-A9 alongside an Altera Cyclone V FPGA as the target.

\subsection{x86 host}

\begin{framed}
\begin{description}
	\item[OS] Ubuntu 12.04 LTS, kernel 3.16
	\item[CPU] Intel Core-i3 ... (two logical core out of four)
	\item[RAM] 1.5GB DDR3
\end{description}
\end{framed}

\subsection{Altera Socrates SoCFPGA}

\begin{framed}
\begin{description}
	\item[OS] Yocto project, kernel 3.14
	\item[CPU] Dual core ARM Cortex-A9, 800MHz
	\item[RAM] ...GB DDR3
	\item[FPGA] Altera Cyclone V
\end{description}
\end{framed}