### 2015-05-04 aknowledgements.tex
conclusion

\addcontentsline{toc}{chapter}{\numberline{}Aknowledgements}
\chapter*{Aknowledgements}

% Talk about the internship? => "Development of verification environment for very high throughput authenticated encryption algorithm"
Thanks a bunch of people here:
- Frédéric Robert for his support
- Sébastien Rabou for his insight
- Batien Heneffe for his extensive help