### 2015-05-09 main.tex
openssl speed dbg on/off

\documentclass[a4paper, 12pt, oneside]{memoir}

\input{preamble}

\begin{document}

\frontmatter

\include{PageDeGarde}
\setcounter{page}{1}

\include{abstract}

\include{aknowledgements}

\selectlanguage{english}
\tableofcontents \pagebreak

\mainmatter

\include{introduction}

\include{technical-background}

\include{presentation}
% TODO merge 'presentation' and 'implementation'?

\include{implementation}

\include{conclusion}

\appendix

\include{toolchain}

\include{cross-compilation}

\include{stuff}

\backmatter

\nocite{*}
\bibliographystyle{plainnat}
\bibliography{bibliography}

\end{document}