### 2015-05-15 introduction.tex

\chapter{Introduction}
% What is hardware offloading?

\section{Challenge}

\section{Network security}