### 2015-05-23 aknowledgements.tex
implementation introduction

\addcontentsline{toc}{chapter}{\numberline{}Aknowledgements}
\chapter*{Aknowledgements}

% Talk about the internship? => "Development of verification environment for very high throughput authenticated encryption algorithm"
Thanks a bunch of people here:

- Frédéric Robert for his support

- Sébastien Rabou for his insight

- Batien Heneffe for his extensive help

Anthony Debruyn for one awesome LaTeX trick.