### 2015-05-24 main.tex
Move presentation to test protocol.

\documentclass[a4paper, 12pt, oneside]{memoir}

\input{preamble}

\begin{document}

\frontmatter

\include{PageDeGarde}
\setcounter{page}{1}
\newgeometry{bottom=2.5cm}

\include{aknowledgements}

\include{abstract}

\selectlanguage{english}
\tableofcontents \pagebreak
\listoffigures \pagebreak

\mainmatter

\include{introduction}

\include{technical-background}

\include{implementation}

\include{test-proto}

\include{results}

\include{conclusion}

\appendix

\include{raw-data}

\include{openssh-patch}

\include{toolchain}

\include{cross-compilation}

\include{stuff}

%\include{configuration}

\backmatter

% \nocite{*}
\bibliographystyle{plainnat}
\bibliography{bibliography}

\end{document}