### 2015-05-24 presentation.tex

\chapter{Presentation}
Here we talk about the protocols, the platform.
Show The OS stack (kernel/user)