### 2015-05-25 aknowledgements.tex
Theory and shit

\addcontentsline{toc}{chapter}{\numberline{}Aknowledgements}
\chapter*{Aknowledgements}

% Talk about the internship? => "Development of verification environment for very high throughput authenticated encryption algorithm"
% Thanks a bunch of people here:

% - Frédéric Robert for his support

% - Sébastien Rabou for his insight

% - Batien Heneffe for his extensive help

% Anthony Debruyn for one awesome LaTeX trick.