### 2015-05-25 main.tex

\documentclass[a4paper, 12pt, oneside]{memoir}

\input{preamble}

\begin{document}

\frontmatter

\include{PageDeGarde}
\setcounter{page}{1}
\newgeometry{bottom=2.5cm}

\include{aknowledgements}

\include{abstract}

\selectlanguage{english}
\tableofcontents \pagebreak
\listoffigures \pagebreak

\mainmatter

\include{introduction}%=> What is the problematic of the thesis? What is the challenge? What do we want to show?

\include{technical-background}%=> Background necessary to understand the problematic and its resolution

\include{implementation}%=> How we will study the problematic, what tool are we going to use to make it clear?

\include{test-proto}%=> How we obtained the results

\include{results}%=> Results. Are they expected? Why are they like that? Analyse.

\include{conclusion}

\appendix

% \include{raw-data}

\include{openssh-patch}

% \include{toolchain}

% \include{cross-compilation}

% \include{stuff}

%\include{configuration}

\backmatter

% \nocite{*}
\bibliographystyle{plainnat}
\bibliography{bibliography}

\end{document}