### 2015-05-26 graphs/openvpn-tls-bench.tex

\begin{tikzpicture}

%%%%%%%%%%%%%%%%%%%%%%%%
% CPU in the background
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        width  = 0.7*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=2pt,%space between the bars
        bar width=16pt,
        enlarge x limits={abs=1},
        ylabel = {CPU\#0 usage},
        hide x axis,
        axis y line*=right,
        ymin=0, ymax=100,
        symbolic x coords={RSA-1024, RSA-2048, RSA-4096},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.15)}, anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]

\addplot[style={black,fill=LimeGreen,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (RSA-1024, 40.32)
        (RSA-2048, 92.14)
        (RSA-4096, 81.97)
    };
    \label{software}

\addplot[style={black,fill=RedOrange,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (RSA-1024, 13.29)
        (RSA-2048, 4.82)
        (RSA-4096, 4.78)
    };
    \label{ba414e}
\legend{software, ba414e}
\end{axis}

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        title = {TLS connections through OpenVPN},
        width  = 0.7*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=10pt,
        bar width=8pt,
        enlarge x limits={abs=1},
        ymajorgrids = true,
        ylabel = {Connections per minute},
        xlabel = {},
        ymin=0,
        symbolic x coords={RSA-1024, RSA-2048, RSA-4096},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.25)}, anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]

\addplot[style={black,fill=ForestGreen,mark=none}]
    coordinates {
        (RSA-1024, 445.4)
        (RSA-2048, 155.6)
        (RSA-4096, 19.6)
    };
    \label{soft-tp}

\addplot[style={black,fill=BrickRed,mark=none}]
    coordinates {
        (RSA-1024, 509.3)
        (RSA-2048, 420.9)
        (RSA-4096, 115.5)
    };
    \label{ba411e-tp}
\legend{}
\end{axis}

\end{tikzpicture}