### 2015-05-28 introduction.tex
Some shit.

\chapter{Introduction}
% What is hardware offloading?

Altera and Xilinx are trying to push those SoC plateforms to the market to popularize the FPGAs.

Power is critical nowadays, esapcially in datacenters.
Several solutions exists: abandonning general purpose processors altoghether and turn towards ARM cores, or add an FPGA to the processor for example.
Intel proposed the latter solution in 2014~\cite{intel-xeon-fpga} and is on its way to expand that market with the merger with Altera~\cite{intel-altera-merger}.

There a re diffirent ways to add cryptographic capabilities to an application:
\begin{itemize}
	\item Having the cipher built in.
	\item Using a cryptographic library, such as OpenSSL.
	\item Using the Linux Crypto API.
\end{itemize}

Present the two IP cores that will be used in this work: BA411E and BA414E.

\section{Challenge}

\section{Network security}