### 2015-05-28 abstract.tex
Comments from Bastien, changed title, added part of the intro.

\thispagestyle{plain}% This way, we can have the page numbering in the foot.
\selectlanguage{english}
\addcontentsline{toc}{chapter}{\numberline{}Abstract}
\begin{abstract}
	\begin{verse}
	\small \textit{
	by Quentin Delhaye,  Master in Computer Science and Engineering, Professional Focus, Université Libre de Bruxelles, 2014--2015.}
	\normalsize
	\end{verse}
	\begin{center}
		\textbf{Implementation of high level cryptographic protocols using a SoC platform}
	\end{center}
This is an abstract. Yep.
\end{abstract}
\clearpage

\thispagestyle{plain}
\selectlanguage{francais}
\addcontentsline{toc}{chapter}{\numberline{}Résumé}
\begin{abstract}
	\begin{verse}
	\small \textit{
	par Quentin Delhaye, Master en ingénieur civil en informatique, à finalité finalité Software and Critical Systems Design, Université Libre de Bruxelles, 2014--2015.}
	\normalsize
	\end{verse}
	\begin{center}
		\textbf{Implementation of high level cryptographic protocols using a SoC platform}
	\end{center}

Abstract, bitch !
\end{abstract}