### 2015-05-29 graphs/ipsec-ftp-bench.tex

\begin{tikzpicture}

%%%%%%%%%%%%%%%%%%%%%%%%
% CPU in the background
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        width  = 0.95*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=2pt,%space between the bars
        bar width=16pt,
        enlarge x limits={abs=1},
        ylabel = {CPU\#0 usage},
        hide x axis,
        axis y line*=right,
        ymin=0, ymax=100,
        symbolic x coords={none:none, aes256cbc:none, aes256cbc:sha256, aes256gcm},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.15)}, anchor=north,legend columns=4},
        enlarge x limits=0.1,
    ]

\addplot[style={black,fill=LimeGreen,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 63.74)
        (aes256cbc:sha256, 74.64)
        (aes256gcm, 89.66)
    };
    \label{software-cpu}

\addplot[style={black,fill=RedOrange,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 14.87)
        (aes256cbc:sha256, 17.25)
        (aes256gcm, 0)
    };
    \label{ba411e-cpu}

\addplot[style={black,fill=gray,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 7.16)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
        (aes256gcm, 0)
    };
    \label{out-of-tunnel-cpu}%"oot" for "out of tunnel"

\addplot[style={black,fill=NavyBlue,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 14.68)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
        (aes256gcm, 0)
    };
    \label{inside tunnel-cpu}%"it" for "in tunnel"
\end{axis}

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        title = {FTP transfer inside an IPSec tunnel},
        width  = 0.95*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=10pt,
        bar width=8pt,
        enlarge x limits={abs=1},
        ymajorgrids = true,
        ylabel = {Throughput [MB/s]},
        xlabel = {},
        ymin=0, ymax=12,
        symbolic x coords={oot, none:none, aes256cbc:none, aes256cbc:sha256, aes256gcm},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.15)}, anchor=north,legend columns=4},
        enlarge x limits=0.1,
    ]

\addplot[style={black,fill=ForestGreen,mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 8.83)
        (aes256cbc:sha256, 6.47)
        (aes256gcm, 5.09)
    };
    \label{software}

\addplot[style={black,fill=BrickRed,mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 8.52)
        (aes256cbc:sha256, 5.80)
        (aes256gcm, 0)
    };
    \label{ba411e}

\addplot[style={black,fill=black,mark=none}]
    coordinates {
        (none:none, 11.39)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
        (aes256gcm, 0)
    };
    \label{out-of-tunnel}%"tp" for "throughput"

\addplot[style={black,fill=MidnightBlue,mark=none}]
    coordinates {
        (none:none, 10.21)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
        (aes256gcm, 0)
    };
    \label{inside tunnel}
\legend{software, ba411e, out-of-tunnel, inside tunnel}
\end{axis}

\end{tikzpicture}