### 2015-05-29 aknowledgements.tex
Aknowledgements

\addcontentsline{toc}{chapter}{\numberline{}Aknowledgements}
\chapter*{Aknowledgements}

% Talk about the internship? => "Development of verification environment for very high throughput authenticated encryption algorithm"
% Thanks a bunch of people here:

% - Frédéric Robert for his support, advices

% - Sébastien Rabou for his insight and the conversations we had on the vast and exciting fields that are semiconductors and security.

% - Batien Heneffe for his extensive help

% Anthony Debruyn for one awesome LaTeX trick.

Before diving into the thesis \textit{per se}, I would like to thank the people who have contributed in a way or an other to this work.\newline{}

\noindent First of all, I am gratefull to Professor Frédéric Robert for his advice and support.\newline{}

\noindent I would also like to thank Sébastien Rabou for his insight and the conversations we had on the vast and exciting fields that are semiconductors and security.\newline{}

\noindent I am also thankful for the extensive help Bastien Heneffe provided for the seven months I spent at Barco Silex office.\newline{}

\noindent Finally, I would like to thank Olivier L'Heureux, Florian Hannesse and Barco Silex team for their help and welcom.\newline{}

\noindent And thank you to Anthony Debruyn for one awesome \LaTeX ~trick.