### 2015-06-21 presentation/ipsec-ftp-cpu.tex
ipsec file transfer

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        % title = {FTP transfer inside an IPSec tunnel},
        width  = \textwidth,
        height = 0.75\paperheight,
        major x tick style = transparent,
        xbar=2pt,
        bar width=8pt,
        % enlarge x limits={abs=1},
        xmajorgrids = true,
        xlabel = {CPU usage},
        ylabel = {},
        % ylabel near ticks,
        % yticklabel pos=right,
        xmin=0, xmax=100,
        symbolic y coords={cipher null, aes:none, aes:sha},
        ytick = data,
        ymajorticks=false,
        scaled x ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.15)}, anchor=north,legend columns=4},
        % enlarge x limits=0.1,
    ]

\addplot[style={black,fill=LimeGreen,postaction={pattern=north east lines},mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (0,cipher null)
        (63.74,aes:none)
        (74.64,aes:sha)
    };
    \label{software-cpu}

\addplot[style={black,fill=gray,postaction={pattern=north east lines},mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (7.16,cipher null)
        (0,aes:none)
        (0,aes:sha)
    };
    \label{out-of-tunnel-cpu}%"oot" for "out of tunnel"

\addplot[style={black,fill=RedOrange,postaction={pattern=north east lines},mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (0,cipher null)
        (14.87,aes:none)
        (17.25,aes:sha)
    };
    \label{ba411e-cpu}

\addplot[style={black,fill=NavyBlue,postaction={pattern=north east lines},mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (14.68,cipher null)
        (0,aes:none)
        (0,aes:sha)
    };
    \label{inside tunnel-cpu}%"it" for "in tunnel"
% \legend{software, ba411e, out-of-tunnel, inside tunnel}
\end{axis}