### 2015-06-21 presentation/ipsec-ftp-throughput.tex

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        % title = {FTP transfer inside an IPSec tunnel},
        width  = \textwidth,
        height = 0.75\paperheight,
        major x tick style = transparent,
        xbar=2pt,
        bar width=8pt,
        % enlarge x limits={abs=1},
        xmajorgrids = true,
        xlabel = {Throughput [MB/s]},
        ylabel = {},
        ylabel near ticks,
        yticklabel pos=right,
        xmin=0, xmax=15,
        symbolic y coords={cipher null, aes:none, aes:sha},
        ytick = data,
        scaled x ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.15)}, anchor=north,legend columns=4},
        % enlarge x limits=0.1,
    ]

\addplot[style={black,fill=ForestGreen,mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (0,cipher null)
        (8.83,aes:none)
        (6.47,aes:sha)
    };
    \label{software}

\addplot[style={black,fill=black,mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (11.39,cipher null)
        (0,aes:none)
        (0,aes:sha)
    };
    \label{out-of-tunnel}%"tp" for "throughput"

\addplot[style={black,fill=BrickRed,mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (0,cipher null)
        (8.52,aes:none)
        (5.80,aes:sha)
    };
    \label{ba411e}

\addplot[style={black,fill=MidnightBlue,mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (10.21,cipher null)
        (0,aes:none)
        (0,aes:sha)
    };
    \label{inside tunnel}
% \legend{software, ba411e, out-of-tunnel, inside tunnel}
\end{axis}